---
title: "Project1"
Author: "Jiayin Liu"
format: html
editor: visual
---

Before starting the project, load all the package that is useful for this project

```{r,echo=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
library(glue)
library(tidycensus)
library(stringr)
library(purrr)
library(tidyr)
```

# Data Processing

## 1.Getting the usual process to work with a given URL and turn the returned info into a nice tibble.

To start any work of data, we need to get the data first. As we want to use API data, it would be nice if the data can be converted to tibble format. To bring in data, we need to build the URL first, and use GET() function to contact the website, then parse the returned data, finally put it into a tibble. Note, it is possible the process failed, thus, it would be safer if we add the option which is 'fail to access the data'.Based on the information from Google,

After get the returned data, I need to make it to a nice tibble format, however, remember that the first row that returned is not the data or the column name. Thus, we need to removed the first row for the output.I also use drop=False to keep the result as a table and not simplify it into a vector. After this step, we already exclude everything we don't want so far.The last step is that give the data the column names. And the we should be good to get teh tibble table data by API now.

```{r}


data_to_tibble <- function(url) {

  returned <- httr::GET(url)

  if (httr::status_code(returned) != 200) {
    stop("FAILED!!!!")
  }

  parsed <- jsonlite::fromJSON(rawToChar(returned$content))

  out <- as_tibble(parsed[-1, , drop = FALSE])

  names(out) <- parsed[1, ]

  out
}



```

## 2. Year of survey (2022 as default). Only a single value here

Based on the requirement, there are basically 2 things I need to do. First, make sure there is only single year value here; second, make sure every year information here is between 2010 to 2022

```{r}
check_year <- function(year = 2022) {
  
  if (length(year) != 1) {
    stop("Please provide only one year.")
  }
  
  if (!year %in% 2010:2022) {
    stop("Year not between 2010 and 2022.")
  }
  
  year
}
```

## 3.Specify the numeric variables to be returned (AGEP and PWGTP as default). PWGTP should always be returned

For the time variables, my strategy is to convert them directly to the minutes in a day. There can be 60\*24=1440 minutes in a day.

```{r}
# create mapping label for use first. Tii many different pattern,, I decide to do this mannuly 
jwdp_map<- c(
 "0"   = NA,
  "001" = 14.5,
  "002" = 44.5,
  "003" = 74.5,
  "004" = 104.5,
  "005" = 134.5,
  "006" = 164.5,
  "007" = 184.5,
  "008" = 194.5,
  "009" = 204.5,
  "010" = 214.5,
  "011" = 224.5,
  "012" = 234.5,
  "013" = 244.5,
  "014" = 254.5,
  "015" = 264.5,
  "016" = 274.5,
  "017" = 284.5,
  "018" = 294.5,
  "019" = 302,
  "020" = 307,
  "021" = 312,
  "022" = 317,
  "023" = 322,
  "024" = 327,
  "025" = 332,
  "026" = 337,
  "027" = 342,
  "028" = 347,
  "029" = 352,
  "030" = 357,
  "031" = 362,
  "032" = 367,
  "033" = 372,
  "034" = 377,
  "035" = 382,
  "036" = 387,
  "037" = 392,
  "038" = 397,
  "039" = 402,
  "040" = 407,
  "041" = 412,
  "042" = 417,
  "043" = 422,
  "044" = 427,
  "045" = 432,
  "046" = 437,
  "047" = 442,
  "048" = 447,
  "049" = 452,
  "050" = 457,
  "051" = 462,
  "052" = 467,
  "053" = 472,
  "054" = 477,
  "055" = 482,
  "056" = 487,
  "057" = 492,
  "058" = 497,
  "059" = 502,
  "060" = 507,
  "061" = 512,
  "062" = 517,
  "063" = 522,
  "064" = 527,
  "065" = 532,
  "066" = 537,
  "067" = 542,
  "068" = 547,
  "069" = 552,
  "070" = 557,
  "071" = 562,
  "072" = 567,
  "073" = 572,
  "074" = 577,
  "075" = 582,
  "076" = 587,
  "077" = 592,
  "078" = 597,
  "079" = 604.5,
  "080" = 614.5,
  "081" = 624.5,
  "082" = 634.5,
  "083" = 644.5,
  "084" = 654.5,
  "085" = 664.5,
  "086" = 674.5,
  "087" = 684.5,
  "088" = 694.5,
  "089" = 704.5,
  "090" = 714.5,
  "091" = 724.5,
  "092" = 734.5,
  "093" = 744.5,
  "094" = 754.5,
  "095" = 764.5,
  "096" = 774.5,
  "097" = 784.5,
  "098" = 794.5,
  "099" = 804.5,
  "100" = 814.5,
  "101" = 824.5,
  "102" = 834.5,
  "103" = 844.5,
  "104" = 854.5,
  "105" = 864.5,
  "106" = 874.5,
  "107" = 884.5,
  "108" = 894.5,
  "109" = 904.5,
  "110" = 914.5,
  "111" = 924.5,
  "112" = 934.5,
  "113" = 944.5,
  "114" = 954.5,
  "115" = 964.5,
  "116" = 974.5,
  "117" = 984.5,
  "118" = 994.5,
  "119" = 1004.5,
  "120" = 1014.5,
  "121" = 1024.5,
  "122" = 1034.5,
  "123" = 1044.5,
  "124" = 1054.5,
  "125" = 1064.5,
  "126" = 1074.5,
  "127" = 1084.5,
  "128" = 1094.5,
  "129" = 1104.5,
  "130" = 1114.5,
  "131" = 1124.5,
  "132" = 1134.5,
  "133" = 1154.5,
  "134" = 1184.5,
  "135" = 1214.5,
  "136" = 1244.5,
  "137" = 1264.5,
  "138" = 1274.5,
  "139" = 1284.5,
  "140" = 1294.5,
  "141" = 1304.5,
  "142" = 1314.5,
  "143" = 1324.5,
  "144" = 1334.5,
  "145" = 1344.5,
  "146" = 1354.5,
  "147" = 1364.5,
  "148" = 1374.5,
  "149" = 1394.5,
  "150" = 1424.5
)
```

```{r}
process_numeric_vars <- function(data, num_vars = c("AGEP", "PWGTP")) {
  
  select_num_vars <- c("AGEP", "GASP", "GRPIP", "JWAP", "JWDP", "JWMNP", "PWGTP")
  
  # Check that all requested numeric variables are allowed
  if (!all(num_vars %in% select_num_vars)) {
    stop("Numeric variables NOT AGEP, GASP, GRPIP, JWAP, JWDP, JWMNP, or PWGTP")
  }
  
  # PWGTP should always be returned
  num_vars <- unique(c(num_vars, "PWGTP"))
  
  # One numeric variable other than PWGTP must be returned (Based on the requirement)
  if (length(setdiff(num_vars, "PWGTP")) < 1) {
    stop("One numeric variable other than PWGTP must be returned")
  }
  
  # For JWAP:
  # JWAP: the number changed in every 5 mins interval, for example 006-->12:25am-12:29 am
  
  convert_jwap <- function(x) {
    
    x <- as.numeric(x)
    
    case_when(
      is.na(x) ~ NA,
      TRUE ~ {
        start_min <- x*5
        mid_min <- start_min + 2
        mid_min %% 1440
      }
    )
  }
  
  #For JWDP, since theere are a lot of different pattern of JWDP, I decide manully insert the mapping information, and this can make the whole process easier
 convert_jwdp <- function(x) {
    
    code <- as.character(x)
    
    code <- ifelse(
      code %in% c("0", "00", "000"),
      "0",
      sprintf("%03d", as.numeric(code))
    )
    
    as.numeric(jwdp_map[code])
  }
  
  for (var in num_vars) {
    
    if (var %in% names(data)) {
      
      if (var == "JWAP") {
        data[[var]] <- convert_jwap(data[[var]])
        
      } else if (var == "JWDP") {
        data[[var]] <- convert_jwdp(data[[var]])
        
      } else {
        data[[var]] <- as.numeric(data[[var]])
      }
    }
  }
  
  data
}
```

## 4. Specify the categorical variables to be returned (SEX as default).

```{r}
process_categorical_vars <- function(data, cat_vars = "SEX") {
  
  selected_cat_vars <- c("FER", "HHL", "SCH", "SCHL", "SEX")
  
  # Check that categorical variables are what we want
  if (!all(cat_vars %in% selected_cat_vars)) {
    stop("Categorical variables NOT FER, HHL, SCH, SCHL, or SEX.")
  }
  
  # One categorical variable must be returned
  if (length(cat_vars) < 1) {
    stop("One categorical variable must be returned")
  }
  
  cat_vars <- unique(cat_vars)
  
  for (var in cat_vars) {
    
    if (var %in% names(data)) {
      
      if (var == "SEX") {
        
        data[[var]] <- factor(
          data[[var]],
          levels = c("1", "2"),
          labels = c("Male", "Female")
        )
        
      } else if (var == "FER") {
        
        data[[var]] <- factor(
          data[[var]],
          levels = c("0", "1", "2"),
          labels = c(
            "N/A (less than 15 years/greater than 50 years/male)",
            "Yes",
            "No"
          )
        )
        
      } else if (var == "HHL") {
        
        data[[var]] <- factor(
          data[[var]],
          levels = c("0", "1", "2", "3", "4", "5"),
          labels = c(
            "N/A (GQ/vacant)",
            "English only",
            "Spanish",
            "Other Indo-European languages",
            "Asian and Pacific Island languages",
            "Other language"
          )
        )
        
      } else if (var == "SCH") {
        
        data[[var]] <- factor(
          data[[var]],
          levels = c("0", "1", "2", "3"),
          labels = c(
            "N/A (less than 3 years old)",
            "No, has not attended in the last 3 months",
            "Yes, public school or public college",
            "Yes, private school or college or home school"
          )
        )
        
      } else if (var == "SCHL") {
        
        data[[var]] <- factor(
          data[[var]],
          levels = c(
            "0", "01", "02", "03", "04", "05", "06", "07", "08", "09",
            "10", "11", "12", "13", "14", "15", "16", "17", "18", "19",
            "20", "21", "22", "23", "24"
          ),
          labels = c(
            "N/A (less than 3 years old)",
            "No schooling completed",
            "Nursery school, preschool",
            "Kindergarten",
            "Grade 1",
            "Grade 2",
            "Grade 3",
            "Grade 4",
            "Grade 5",
            "Grade 6",
            "Grade 7",
            "Grade 8",
            "Grade 9",
            "Grade 10",
            "Grade 11",
            "12th grade - no diploma",
            "Regular high school diploma",
            "GED or alternative credential",
            "Some college, but less than 1 year",
            "1 or more years of college credit, no degree",
            "Associate's degree",
            "Bachelor's degree",
            "Master's degree",
            "Professional degree beyond a bachelor's degree",
            "Doctorate degree"
          )
        )
      }
    }
  }
  
  data
}
```

## 5.Specify the geography level: Region, Division, or State (with the default geograph level of State).

Since there are a great number of information in this geography level, I also will mannully enter the categories.

```{r}
process_geography <- function(data, geography = "ST") {
  
  geography <- toupper(geography)
  
  selected_geography <- c("ST", "DIVISION", "REGION")
  
  # Check if the geography variable is what we want
  if (!(geography %in% selected_geography)) {
    stop("Geography must be one of: ST, DIVISION, or REGION.")
  }
  
  # State labels
  state_levels <- c(
    "01", "02", "04", "05", "06", "08", "09", "10", "11", "12",
    "13", "15", "16", "17", "18", "19", "20", "21", "22", "23",
    "24", "25", "26", "27", "28", "29", "30", "31", "32", "33",
    "34", "35", "36", "37", "38", "39", "40", "41", "42", "44",
    "45", "46", "47", "48", "49", "50", "51", "53", "54", "55",
    "56", "72"
  )
  
  state_labels <- c(
    "Alabama/AL", "Alaska/AK", "Arizona/AZ", "Arkansas/AR",
    "California/CA", "Colorado/CO", "Connecticut/CT", "Delaware/DE",
    "District of Columbia/DC", "Florida/FL", "Georgia/GA", "Hawaii/HI",
    "Idaho/ID", "Illinois/IL", "Indiana/IN", "Iowa/IA", "Kansas/KS",
    "Kentucky/KY", "Louisiana/LA", "Maine/ME", "Maryland/MD",
    "Massachusetts/MA", "Michigan/MI", "Minnesota/MN", "Mississippi/MS",
    "Missouri/MO", "Montana/MT", "Nebraska/NE", "Nevada/NV",
    "New Hampshire/NH", "New Jersey/NJ", "New Mexico/NM", "New York/NY",
    "North Carolina/NC", "North Dakota/ND", "Ohio/OH", "Oklahoma/OK",
    "Oregon/OR", "Pennsylvania/PA", "Rhode Island/RI",
    "South Carolina/SC", "South Dakota/SD", "Tennessee/TN", "Texas/TX",
    "Utah/UT", "Vermont/VT", "Virginia/VA", "Washington/WA",
    "West Virginia/WV", "Wisconsin/WI", "Wyoming/WY", "Puerto Rico/PR"
  )
  
  # Division labels
  division_levels <- c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9")
  
  division_labels <- c(
    "Puerto Rico",
    "New England (Northeast region)",
    "Middle Atlantic (Northeast region)",
    "East North Central (Midwest region)",
    "West North Central (Midwest region)",
    "South Atlantic (South region)",
    "East South Central (South region)",
    "West South Central (South region)",
    "Mountain (West region)",
    "Pacific (West region)"
  )
  
  # Region labels
  region_levels <- c("1", "2", "3", "4", "9")
  
  region_labels <- c(
    "Northeast",
    "Midwest",
    "South",
    "West",
    "Puerto Rico"
  )
  
  # Convert selected geography variable
  if (geography %in% names(data)) {
    
    if (geography == "ST") {
      
      data[[geography]] <- factor(
        data[[geography]],
        levels = state_levels,
        labels = state_labels
      )
      
    } else if (geography == "DIVISION") {
      
      data[[geography]] <- factor(
        data[[geography]],
        levels = division_levels,
        labels = division_labels
      )
      
    } else if (geography == "REGION") {
      
      data[[geography]] <- factor(
        data[[geography]],
        levels = region_levels,
        labels = region_labels
      )
    }
  }
  
  data
}
```

## 6. Subset the data

```{r}
# I tried to use ST, DIVISITON, and REGION directly, but the API doesnt allowed me to use these variables directly as predicates, so we use ucgid for API-level subsetting.
process_geo_subset <- function(geography = "ST", geo_values = "24") {
  
  geography <- toupper(geography)
  
  allowed_geography <- c("ST", "DIVISION", "REGION")
  
  if (!(geography %in% allowed_geography)) {
    stop("Geography must be one of: ST, DIVISION, or REGION.")
  }
  
  # If user asks for all values
  if (length(geo_values) == 1 && tolower(geo_values) == "all") {
    
    if (geography == "ST") {
      return(list(ucgid = "040C0US*"))
    }
    
    if (geography == "DIVISION") {
      return(list(ucgid = "030C0US*"))
    }
    
    if (geography == "REGION") {
      return(list(ucgid = "020C0US*"))
    }
  }
  
  geo_values <- as.character(geo_values)
  
  # Add leading zero for one-digit state codes
  if (geography == "ST") {
    
    geo_values <- ifelse(
      nchar(geo_values) == 1,
      paste0("0", geo_values),
      geo_values
    )
    
    ucgid_values <- paste0("040C0US", geo_values)
  }
  
  if (geography == "DIVISION") {
    ucgid_values <- paste0("030C0US", geo_values)
  }
  
  if (geography == "REGION") {
    ucgid_values <- paste0("020C0US", geo_values)
  }
  
  list(
    ucgid = paste(ucgid_values, collapse = ",")
  )
}
  

  
```

## 7. Apply these small function into 1 main function for query

I put all the default information first

```{r}

get_pums_data <- function(year = 2022,
                          num_vars = c("AGEP", "PWGTP"),
                          cat_vars = "SEX",
                          geography = "ST",
                          geo_values = "24",
                          key) {
  
  # Check year
  year <- check_year(year)
  
  # Check geography
  geography <- toupper(geography)
  
  selected_geography <- c("ST", "DIVISION", "REGION")
  
  if (!(geography %in% selected_geography)) {
    stop("Geography must be one of: ST, DIVISION, or REGION.")
  }
  
  # Check numeric variables
  selected_num_vars <- c(
    "AGEP", "GASP", "GRPIP", "JWAP", "JWDP", "JWMNP", "PWGTP"
  )
  
  if (!all(num_vars %in% selected_num_vars)) {
    stop("Numeric variables must be AGEP, GASP, GRPIP, JWAP, JWDP, JWMNP, or PWGTP.")
  }
  
  # PWGTP must always be returned
  num_vars <- unique(c(num_vars, "PWGTP"))
  
  if (length(setdiff(num_vars, "PWGTP")) < 1) {
    stop("At least one numeric variable other than PWGTP must be returned.")
  }
  
  # Check categorical variables
  selected_cat_vars <- c("FER", "HHL", "SCH", "SCHL", "SEX")
  
  if (!all(cat_vars %in% selected_cat_vars)) {
    stop("Categorical variables must be FER, HHL, SCH, SCHL, or SEX.")
  }
  
  if (length(cat_vars) < 1) {
    stop("At least one categorical variable must be returned.")
  }
  
  # Variables to return
  vars <- unique(c(num_vars, cat_vars, geography))
  
  query_list <- list(
    get = paste(vars, collapse = ","),
    key = as.character(key)
  )
  
  # Add geography subset using ucgid
  geo_query <- process_geo_subset(
    geography = geography,
    geo_values = geo_values
  )
  
  query_list <- c(query_list, geo_query)
  
  # API URL
  api_url <- paste0(
    "https://api.census.gov/data/",
    year,
    "/acs/acs1/pums"
  )
  
  # API call
  res <- httr::GET(
    url = api_url,
    query = query_list
  )
  
  # Read API result
  text <- rawToChar(res$content)
  
  if (!startsWith(text, "[")) {
    stop(text)
  }
  
  parsed <- jsonlite::fromJSON(text)
  
  # Header row
  header_names <- parsed[1, ]
  
  # Make duplicated names unique
  header_names <- make.unique(header_names)
  
  # Data rows
  out <- tibble::as_tibble(
    parsed[-1, , drop = FALSE],
    .name_repair = "unique"
  )
  
  names(out) <- header_names
  
  # If duplicated geography columns were created, keep the original one
  # This handles names like ST and ST.1
  duplicate_geo_cols <- grep(
    paste0("^", geography, "\\."),
    names(out),
    value = TRUE
  )
  
  if (length(duplicate_geo_cols) > 0) {
    out <- dplyr::select(out, -dplyr::all_of(duplicate_geo_cols))
  }
  
  # Process returned variables
  out <- process_numeric_vars(
    data = out,
    num_vars = num_vars
  )
  
  out <- process_categorical_vars(
    data = out,
    cat_vars = cat_vars
  )
  
  out <- process_geography(
    data = out,
    geography = geography
  )
  
  # Add year variable
  out$year <- year
  
  # Add census class
  class(out) <- c("census", class(out))
  
  out
}

#Test for the data, if I only want data from 2022 and only want the data from Maryland 
my_key <- "8ed6237bac2c1c1094124f4d196b717949e7fb18"

data <- get_pums_data(
  year = 2022,
  num_vars = c("AGEP", "PWGTP",'JWAP','JWDP'),
  cat_vars = c("SEX","SCHL"),
  geography = "ST",
  geo_values = "24",
  key = my_key
)

head(data)# Yes! This works!

```

## 8. write a function let people specify multiple years of survey data

```{r}
get_pums_multi_year <- function(years,
                                num_vars = c("AGEP", "PWGTP"),
                                cat_vars = "SEX",
                                geography = "ST",
                                geo_values = "24",
                                key) {
  
  # Check years like the first function 
  if (!is.numeric(years)) {
    stop("Years should be numeric.")
  }
  
  if (!all(years %in% 2010:2022)) {
    stop("All years between 2010 and 2022.")
  }
  
  # Do not allow 2020
  if (2020 %in% years) {
    stop("Do not include 2020 due to COVID.")
  }
  
  # get the data for each year separately and combine this with the function i wrote before
  out <- purrr::map_dfr(
    years,
    ~ get_pums_data(
      year = .x,
      num_vars = num_vars,
      cat_vars = cat_vars,
      geography = geography,
      geo_values = geo_values,
      key = key
    )
  )
  
  # Add census class
  class(out) <- c("census", class(out))
  
  out
}

# check if the code works well 
multi_data <- get_pums_multi_year(
  years = c(2019, 2021, 2022),
  num_vars = c("AGEP", "PWGTP","GRPIP"),
  cat_vars = c("SEX","SCH","SCHL"),
  geography = "ST",
  geo_values = "24",
  key = my_key
)

head(multi_data)#YES! works!
```

# Obtain person level record

```{r}
get_person_level_records <- function(data) {
  
  # Check PWGTP exists and make it as numeric
  if (!("PWGTP" %in% names(data))) {
    stop("PWGTP mustin the data")
  }
  
  data$PWGTP <- as.numeric(data$PWGTP)
  
  # Check for missing 
  if (any(is.na(data$PWGTP))) {
    stop("PWGTP missing")
  }
  
  if (any(data$PWGTP < 0)) {
    stop("PWGTP not reasonable")
  }
  
  # replicate the data
  person_data <- data %>%
    tidyr::uncount(weights = PWGTP,.remove=FALSE)
  
  person_data
}

#check if the function works well
rep_dt<-get_person_level_records(multi_data)#works well!
```

# Write a function for summary

For summary information, I personally add another option -- group_by. Since the previous reuquirement give me the option for year, I want to see the effect of year in the summary information. In that case, grouo_by would be very helpful to provide more information.

## 1. for summary

```{r}
summary_census <- function(object,
                           num_vars = NULL,
                           cat_vars = NULL,
                           group_var = NULL,
                           ...) {
  
  # Check PWGTP
  if (!("PWGTP" %in% names(object))) {
    stop("PWGTP must in the data")
  }
  
  object$PWGTP <- as.numeric(object$PWGTP)
  
  # check group variable
  if (!is.null(group_var)) {
    
    if (!(group_var %in% names(object))) {
      stop("group_var not in the data.")
    }
  }
  

    # numeric variables:
  if (is.null(num_vars)) {
    
    num_vars <- names(object)[sapply(object, is.numeric)]
    #remove the variable I don't want to summarize,not useful
    num_vars <- setdiff(num_vars,c("PWGTP", "JWAP", "JWDP"))
  }
  
  # categorical variables:
  if (is.null(cat_vars)) {
    cat_vars <- names(object)[sapply(object, is.factor)]
  }
  
  # Check numeric variables
  if (!all(num_vars %in% names(object))) {
    stop("Some numeric variables not in the data.")
  }
  
  # Check categorical variables
  if (!all(cat_vars %in% names(object))) {
    stop("Some categorical variables  not in the data.")
  }
  
  # Numeric summaries: weighted mean and weighted SD
 numeric_summary <- lapply(num_vars, function(var) {
    
    temp_data <- object %>%
      filter(
        !is.na(.data[[var]]),
        !is.na(PWGTP)
      )
    
    if (!is.null(group_var)) {
      
      temp_data %>%
        group_by(.data[[group_var]]) %>%
        summarise(
          weighted_mean = sum(as.numeric(.data[[var]]) * PWGTP) / sum(PWGTP),
          weighted_sd = sqrt(
            sum((as.numeric(.data[[var]])^2) * PWGTP) / sum(PWGTP) -
              weighted_mean^2
          ),
          .groups = "drop"
        ) %>%
        rename(group = 1) %>%
        mutate(variable = var) %>%
        select(variable, group, weighted_mean, weighted_sd)
      
    } else {
      
      x <- as.numeric(temp_data[[var]])
      w <- temp_data$PWGTP
      
      weighted_mean <- sum(x * w) / sum(w)
      
      weighted_sd <- sqrt(
        sum((x^2) * w) / sum(w) - weighted_mean^2
      )
      
      tibble(
        variable = var,
        weighted_mean = weighted_mean,
        weighted_sd = weighted_sd
      )
    }
  })
  
  names(numeric_summary) <- num_vars
  
  categorical_summary <- lapply(cat_vars, function(var) {
    
    temp_data <- object %>%
      filter(
        !is.na(.data[[var]]),
        !is.na(PWGTP)
      ) %>%
      mutate(level = .data[[var]])
    
    if (!is.null(group_var)) {
      
      temp_data %>%
        group_by(.data[[group_var]], level) %>%
        summarise(
          weighted_count = sum(PWGTP, na.rm = TRUE),
          .groups = "drop"
        ) %>%
        mutate( percent = weighted_count / sum(weighted_count) * 100) %>% 
        rename(group = 1) %>%
        mutate(variable = var) %>%
        select(variable, group, level, weighted_count,percent)
      
    } else {
      
      temp_data %>%
        group_by(level) %>%
        summarise(
          weighted_count = sum(PWGTP, na.rm = TRUE),
          .groups = "drop"
        ) %>%
        mutate(variable = var) %>%
        select(variable, level, weighted_count)
    }
  })
  
  names(categorical_summary) <- cat_vars
  
  out <- list(
    numeric_summary = numeric_summary,
    categorical_summary = categorical_summary
  )
  
  out
}

#test it 
summary_census(
  rep_dt,
  num_vars = c("AGEP","GRPIP"),
  cat_vars = c("SEX","SCH","SCHL"),
  group_var="year"
)->a

a$numeric_summary
a$categorical_summary
```

## 2. for plot

```{r}

class(rep_dt) <- c("census", class(rep_dt))

#Plot function 
plot_census <- function(x, cat_var, num_var, ...) {
  # check census class
  if (!inherits(x, "census")) {
    stop("x must have class 'census'")
  }
  
  # Check PWGTP
  if (!("PWGTP" %in% names(x))) {
    stop("PWGTP must be in the data.")
  }
 
  # Check categorical variables
  if (!all(cat_var %in% names(x))) {
    missing_cat <- cat_var[!(cat_var %in% names(x))]
    stop(paste("The categorical variable not in the data:",
               paste(missing_cat, collapse = ", ")))
  }
 
  # Check numeric variables
  if (!all(num_var %in% names(x))) {
    missing_num <- num_var[!(num_var %in% names(x))]
    stop(paste("The numeric variable not in the data:",
               paste(missing_num, collapse = ", ")))
  }
 
  # Make sure weight is numeric
  x$PWGTP <- as.numeric(x$PWGTP)
 
  # Store plots
  plot_list <- list()
 
  # Create one plot for each cat_var and num_var combination
  for (cat in cat_var) {
    for (num in num_var) {
     
      x[[num]] <- as.numeric(x[[num]])
     
      p <- ggplot(
        data = x,
        aes(x = .data[[cat]],y = .data[[num]],weight = PWGTP)) +
        geom_boxplot() +
        labs(x = cat,y = num,title = paste("Weighted boxplot of", num, "by", cat)
        )
     
      plot_list[[paste(cat, num, sep = "_")]] <- p
    }
  }
 
  return(plot_list)
}




#This is for the example to run faster, and the prebious function can run different combination, just takes tooooo long to run. (For me over 1 hour)
plot_faster<-function(x, cat_var, num_var, max_n = 5000, ...) {
  
  # check census class
  if (!inherits(x, "census")) {
    stop("x must have class 'census'")
  }
  
  # check PWGTP
  if (!("PWGTP" %in% names(x))) {
    stop("PWGTP must be in the data.")
  }
  
  # categorical variable
  if (length(cat_var) != 1) {
    stop("one categorical variable.")
  }
  
  # numeric variable
  if (length(num_var) != 1) {
    stop(" one numeric variable.")
  }
  
  # check selected variables
  if (!(cat_var %in% names(x))) {
    stop("The categorical variable not in the data.")
  }
  
  if (!(num_var %in% names(x))) {
    stop("The numeric variable not in the data.")
  }
  
  # keep only needed columns to make it faster
  plot_data <- x %>% 
    select(all_of(c(cat_var, num_var, "PWGTP"))) %>% 
    filter(
      !is.na(.data[[cat_var]]),
      !is.na(.data[[num_var]]),
      !is.na(PWGTP)
    )
  
  # sample rows for faster plotting
  if (nrow(plot_data) > max_n) {
    set.seed(123)
    plot_data <- plot_data %>% 
      slice_sample(n = max_n)
  }
  
  # make sure variables are correct type
  plot_data[[cat_var]] <- as.factor(plot_data[[cat_var]])
  plot_data[[num_var]] <- as.numeric(plot_data[[num_var]])
  plot_data$PWGTP <- as.numeric(plot_data$PWGTP)
  
  # boxplot
  ggplot(
    data = plot_data,
    aes(
      x = get(cat_var),
      y = get(num_var),
      weight = PWGTP
    )
  ) +
    geom_boxplot(...) +
    labs(
      x = cat_var,
      y = num_var,
      title = paste("Weighted boxplot of", num_var, "by", cat_var)
    ) +
    theme_minimal()
}

#test for the function
plot_faster(
  rep_dt,
  cat_var = "SEX",
  num_var = "AGEP"
)
```

# Summary

In the project instruction, it mentions we want to exclude year 2020 due to COVID-19. This drive my interests that how COVID-19 affects these factors includes Gross rent as a percentage of household income past 12 months, Educational attainment situation, and School enrollment situation. 

The weighted mean of AGEP (age) was 35.9 in 2019, 36.3 in 2021, and 36.2 in 2022, with standard deviations around 22 years in all three years. This shows that the overall age distribution did not change a lot before and after COVID. I think this is good, since this makes other comparison more strong. For GRPIP, the weighted mean decreased from 15.0 in 2019 to 14.1 in 2021 and then to 13.5 in 2022. in one word, the gross rent as a percentage of household income past 12 months keep droping. This indicates that the average gross rent percentage of income was lower in the post-COVID years in this sample. Although the standard deviations were still large, around 24 to 25, this can cause by a lot of variation among individuals or households. In addition, it is hard to hard to judge if this means families are living better life or not, since this dropping percentage can caused by moving to lower rent home, or increasing daily life budge push the families.

For the categorical variables, the weighted counts for SEX show that females were slightly more represented than males in each year. In 2019, the weighted count was about 510.4 million for males and 523.5 million for females. But in 2021, there was about 484.6 million for males and 501.3 million for females. And when it comes to 2022, the number increased to about 523.7 million for males and 535.8 million for females. For school enrollment (SCH), the largest category in every year was “No, has not attended in the last 3 months,” and the scond is “Yes, public school or public college.” This school enrollment pattern was consistent before and after COVID. For education attainment, In 2019 Regular high school diploma has the highest percentatge followed by bachelor's degree. and this is the same thing for 2021 and 2022, but the percentage of having bachelor's degree is increasing while the percentage of regular high school diploma is decreasing. Overall, the results told us that age and sex composition were quite stable across the years, while GRPIP showed a clearer decline after COVID. The school enrollment categories also showed similar thing across years. there was a small decrease in public school or public college enrollment in 2021 ,and then an increase again in 2022. People seems like to stay more in school for higher education after the COVID.